# Exploratory Data Analysis

EDA(Exploratory Data Analysis) is necessary for data analyst to check the dataset before statistical modeling(or machine learning)

## head and tail

```{r}
df_mpg <- data.frame(ggplot2::mpg)
head(df_mpg)
```

we have loaded `mpg` dataset from `ggplot2` package. `ggplot2` contains multiple useful datasets and various visualization tools.
`head` function extract top 6 rows of input dataset. In contrast, we can extract bottom 6 rows using `tail` function.

```{r}
tail(df_mpg)
```

It's important to check rows in this way, because most of dataset we meet are so large that it can't be shown just in one page.

## dim and str

using `dim` function, we can check the size of dataset. The size means rows x columns.

```{r}
dim(df_mpg)
```

This `mpg` dataset has 234 rows and 11 columns(features).

We do not need to use this function just for check entire dataset.

```{r}
dim(head(df_mpg))
```

```{r}
dim(df_mpg[,2:4])
```
```{r}
dim(df_mpg['model'])
```

but we can't use `dim` for 1-dim vector. Note that df_mpg['model'] is not a vector, it's dataframe yet.
```{r}
str(df_mpg['model'])
```

It's different from df_mpg$model
```{r}
str(df_mpg$model)
```
This is vector. And it's same with df_mpg[['model']]
```{r}
str(df_mpg[['model']])
```


## summary

we can check summary statistics using `summary` function.

```{r}
summary(df_mpg)
```


If column is categorical variable, like `model`, `trans`, `drv`, `manufacturer`, `fl`, `class`, it just shows Length, Class and Mode. But in case of numerical, it can shows mean of variable, median of variable, minimum of variable, maximum of variable and 1st and 3rd quartile.


## table

And R also provides a frequency table. We can use this as call `table` function.
```{r}
table(df_mpg$model)
```

```{r}
table(df_mpg$class)
```

It's helpful to check counts of each values in variable. To better use, we can use this table with visualizing original variable.

first, let's import `ggplot2` library
```{r}
library(ggplot2)
```


```{r}
qplot(df_mpg$class)
```
You should be careful to visualize the table as it is. If you visualize the table, it shows the different one.

```{r}
qplot(table(df_mpg$class))
```

## qplot

Visualization is always important part for EDA. Let's study more about `qplot`.
we use `mpg` dataset. Extract top 5 rows
```{r}
head(mpg)
```

`qplot` means "quick plot". It is useful for quickly running our plot.

let's see some feature's frequency using `qplot`

```{r}
qplot(data=mpg, x=manufacturer)
```

```{r}
qplot(data=mpg, x=fl)
```

```{r}
qplot(data=mpg, x=year)
```

it's weired. `year` data have only 2 cases.
let's check that.

```{r}
table(mpg$year)
```

yes. it's contain only the car data that manufacturered in 1999 and 2008. We have to think about this when analysis this dataset.

and let's go ahead and check the other feature one more.

```{r}
qplot(data=mpg, x=class)
```

like this, we can show data's frequency by input x(feature). 

and `class` contains types of car like size or shape etc.
```{r}
table(mpg$class)
```


Then, what will happen if we input y into qplot function?

```{r}
qplot(data=mpg, x=class, y=hwy)
```

This is scatter plot. From this plot, we can check highway fuel economy of each car class. Pick-up truck and SUV are not good at average fuel economy.

But, we have more useful plot than scatter plot for checking the mean of data. Let's use that.

```{r}
qplot(data=mpg, x=class, y=hwy, geom='boxplot', color=class)
```

It's box plot. We can check means easily and even outliers. It's very important that how our data visualized and what tools we use.

If we use line plot for this situation, It is difficult to accurately show the data.

```{r}
qplot(data=mpg, x=class, y=hwy, geom='line', color=class)
```


## hist

R have basic histplot function. We can use this by calling `hist()`. `hist` means histogram. It shows the distribution of continuous(numeric) variable.

```{r}
hist(mpg$hwy)
```

If you input categorical variable(ex. mpg$class) into `hist` function, 

it returns error that `Error in hist.default(mpg$class) : 'x' must be numeric`





