# Data Visualization

## qplot


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


first, let's import `ggplot2` library
```{r}
library(ggplot2)
```

and we use `mpg` dataset. Extract top 5 rows
```{r}
head(mpg)
```

`qplot` means "quick plot". It is useful for quickly running our plot.

let's see some feature's frequency using `qplot`

```{r}
qplot(data=mpg, x=manufacturer)
```

```{r}
qplot(data=mpg, x=fl)
```

```{r}
qplot(data=mpg, x=year)
```

it's weired. `year` data have only 2 cases.
let's check that.

```{r}
table(mpg$year)
```

yes. it's contain only the car data that manufacturered in 1999 and 2008. We have to think about this when analysis this dataset.

and let's go ahead and check the other feature one more.

```{r}
qplot(data=mpg, x=class)
```

like this, we can show data's frequency by input x(feature). 

and `class` contains types of car like size or shape etc.
```{r}
table(mpg$class)
```


Then, what will happen if we input y into qplot function?

```{r}
qplot(data=mpg, x=class, y=hwy)
```

This is scatter plot. From this plot, we can check highway fuel economy of each car class. Pick-up truck and SUV are not good at average fuel economy.

But, we have more useful plot than scatter plot for checking the mean of data. Let's use that.

```{r}
qplot(data=mpg, x=class, y=hwy, geom='boxplot', color=class)
```

It's box plot. We can check means easily and even outliers. It's very important that how our data visualized and what tools we use.

If we use line plot for this situation, It is difficult to accurately show the data.

```{r}
qplot(data=mpg, x=class, y=hwy, geom='line', color=class)
```


## hist

`hist` means histogram. It shows the distribution of continuous(numeric) variable.

```{r}
hist(mpg$hwy)
```

If you input categorical variable(ex. mpg$class) into `hist` function, 

it returns error that `Error in hist.default(mpg$class) : 'x' must be numeric`




