# Data Preprocessing

While exploring data, we need to preprocess it properly for our purpose.
And these work can be repeat until we construct a dataset that can make our probabilistic model well.

## ifelse

`ifelse` is `Conditional Statments function` that is very simple and useful in R. We can use this with just 3 parameters - condition, return value when it's true, return value when it's false.

let's see example code.

first, print the head of dataset
```{r}
head(mpg$hwy)
```

And let's apply ifelse to this head.
if the hwy(highway fuel efficiency) is higher than 30, returns 'Good'.
```{r}
ifelse(head(mpg$hwy) > 30, 'Good', 'Bad')
```

We can use this data by assign method(`<-`). Let's Control Entire variable.

```{r}
evaluate_hwy <- ifelse(mpg$hwy > 30, 'Good', 'Bad')
print(evaluate_hwy)
```

Then, Why is this necessary?

Sometimes we need categorical data(nominal data or ordinal data) for not only visualization but also data handling. Especially, We do not always get gaussian distribution dataset, so since the data is not uniform, it is necessary to divide it into appropriate intervals.

We can use this function to not only create new variable but also add another column.
```{r}
df_mpg <- mpg
df_mpg['eval_hwy'] <- ifelse(df_mpg$hwy>30, 'High', 'Low')
df_mpg$eval_hwy
```

## dplyr :: mutate

In R, `dplyr` package is the most popular library for data preprocessing. Especially it provides pipe('%>%'), and it's very helpful for intuitive coding. And `mutate` in `dplyr` package do exactly same work as the one above, However it is simpler and easier to use. Let's see this.

(before load, you have to install "dplyr" package)
```{r}
# install.packages("dplyr") 
library(dplyr)
```

```{r}
head(df_mpg)
```
We'll use `mutate` function and `pipe` syntex to handle `cty`(city fuel economy) column.

```{r}
df_mpg %>%
  mutate(mean_fuel = (cty + hwy)/2) %>%
  head
```

We can see the last column is added and named with `mean_fuel`. We didn't have to use the name of dataframa(df_mpg) repeatly to use columns. It's very nice.

And of course we can use `ifelse` function with the `pipe` and `mutate`.

```{r}
df_mpg %>%
  mutate(mean_fuel = (cty + hwy)/2, 
         eval_fuel = ifelse(mean_fuel > 20, 'high', 'low')) %>%
  head
```

If you want to use this data table continuously, you have to assign this table to the new variable like this.

```{r}
new_mpg <- df_mpg %>%
            mutate(mean_fuel = (cty + hwy)/2, 
            eval_fuel = ifelse(mean_fuel > 20, 'high', 'low'))
```

```{r}
str(new_mpg)
```

```{r}
qplot(new_mpg$mean_fuel)
```

```{r}
qplot(new_mpg$eval_fuel)
```

We can use more functions from `dplyr` package. The commonly used functions that except mutate we have seen already are `filter()`, `select()`, `arrange()`, `summarise()`. Let's look at these one by one.

## dplyr :: filter

First, `filter()` function. `filter` extracts rows by some conditions. 
```{r}
head(new_mpg)
```
If you want to extract only `audi`, you can use `filter`.

```{r}
new_mpg %>%
  filter(manufacturer=='audi')
```

If you want to extract `audi a4` model, 
```{r}
new_mpg %>%
  filter(manufacturer=='audi' & model=='a4')
```

If you want to extract `audi a4` that have `Good for highway fuel economy`,
```{r}
new_mpg %>%
  filter(manufacturer=='audi' & model=='a4' & eval_hwy == 'High')
```
Yes, now you can choose the car that you want.

## dplyr :: select

next, `select()` function.

